---
title: "day1-points"
author: "@jaanekaraster"
date: "2022-11-01"
output: html_document
---

## The Great Potoo: An Alien Bird
The Great Potoo (*Nyctibius grandis*) is a weird bird. It's the subject of legends (https://cuyabenolodge.com.ec/tell-me-the-legend-of-the-singing-bird-potoo/) due to its haunting cry. 


It looks rather helpless and vaguely demonic. It's also hard to find in the jungle because it blends into the tree bark. 

Nevertheless, its predators include spider monkeys (Ateles geoffroyi)

Potoo data: GBIF.org (29 October 2022) GBIF Occurrence Download https://doi.org/10.15468/dl.wnhjbz

Spider monkey data: GBIF.org (01 November 2022) GBIF Occurrence Download https://doi.org/10.15468/dl.wesxm7

- Panamanian white-faced capuchin: GBIF.org (01 November 2022) GBIF Occurrence Download https://doi.org/10.15468/dl.545r53




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r libraries}
#install.packages(c("cowplot", "googleway", "ggplot2", "ggrepel", 
#"ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata"))

```

```{r get-data}
#potoo data
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(tidyr) # to clean up NA's
library(ggplot2)
library(jpeg)
library(patchwork)
library(extrafont) # for new fonts
library(remotes) # for installing fonts
#remotes::install_version("Rttf2pt1", version = "1.3.8")
#pkgbuild::check_build_tools(debug=TRUE)
#extrafont::font_import()

#font_import()
#fonttable()
path = "potoo_wikipedia.jpg"
img = readJPEG(path, native = TRUE)

#read in the CSV's
#potoo data
potoo_csv = read.csv("potoo_data.csv")
potoo_data = st_as_sf(potoo_csv, coords=c("decimalLongitude","decimalLatitude"), crs=4326)
#spider monkey data
monkey_csv = read.csv("spider_monkey.csv") %>% drop_na(decimalLongitude)
monkey_data = st_as_sf(monkey_csv, coords=c("decimalLongitude","decimalLatitude"), crs=4326)
#capuchin data
capuchin_csv = read.csv("capuchin.csv",) %>% drop_na(decimalLongitude)
capuchin_data = st_as_sf(monkey_csv, coords=c("decimalLongitude","decimalLatitude"), crs=4326)

#get natural earth layers
world <- ne_countries(scale='medium',returnclass = 'sf')
#urban_areas = ne_download(scale = 50, type = 'urban_areas', category = 'cultural',returnclass = 'sf')
#pop = ne_download(scale = 110, type = 'populated_places', category = 'cultural',returnclass = 'sf')
my_y_title = expression(paste(italic("Nyctibius grandis"), ": An Alien Bird"))
#plot the map
(base<- ggplot(data = world) +
     geom_sf(fill = "antiquewhite3") + 
  #geom_sf(data = urban_areas, colour = "black")+
  #geom_sf(data = pop, colour = "black") + 
    geom_sf(data = potoo_data, colour = "red", size = 0.01) +
    annotate(geom = "text", x = -75, y = 15, label = "Caribbean Sea", 
    fontface = "italic", color = "grey22", size = 3) +
    annotate(geom = "text", x = -90, y = 5, label = "Pacific Ocean", 
    fontface = "italic", color = "grey22", size = 3) +
    coord_sf(xlim = c(-95, -65), ylim = c(20,-5), expand = FALSE) +
    theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold", color="black"), plot.subtitle = element_text(hjust = 0.5), legend.position = "none", panel.background = element_rect(fill = "antiquewhite1"),
     panel.border = element_rect(fill = NA))) + 
  labs(x = "", y = "", title = "Realm of the Great Potoo", subtitle = my_y_title, caption = "@jaanekaraster | Data: GBIF.org, Natural Earth") +  ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey30", "white"),
    text_family = "Maiandra GD"
  ) +
  ggspatial::annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.2, "in"), pad_y = unit(0.4, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "Arial"
    )
  ) +
   inset_element(p = img,
                left = 0,
                bottom = 0.7,
                right = 0.2,
                top = 1,clip = TRUE) + theme(rect = element_rect(fill = "transparent"))

ggsave("potoo_range.png")
```
